Precise predictions for same-sign W-bosons scattering at the LHC

### 2018-02-15 Draft/abstract.tex
further small changes

  Vector-boson scattering processes are of great importance for the
  current run-2 and future runs at design energy of the Large Hadron
  Collider. The presence of triple and quartic gauge couplings in the process
  gives access the gauge sector of the Standard Model and possible new-physics
  contributions there. 
  To test this hypothesis, a precise knowledge of the Standard
  Model contributions is necessary, 
  with a precision which at least matches the
  experimental uncertainties
  of existing and forthcoming measurements.
  In this article, we present a detailed study of the vector-boson
  scattering process with two positively-charged leptons and missing
  transverse momentum in the final state, mediated predominantly by
  same-sign production of two W bosons with positive charge.
  In particular, we first carry out a systematic comparison of the various 
  approximations that are usually performed for this kind of process against the complete calculation, 
  at LO and NLO QCD accuracy. Such  a study is performed both 
  in the usual fiducial region used by experimental collaborations and in a more inclusive phase space,
  where the differences among the various approximations lead to more sizeable effects.
  Afterwards, we turn ourselves to predictions matched to parton showers, at LO and NLO: we show that on the
  one hand, the inclusion of NLO QCD corrections leads to more stable predictions, but on the other the details 
  of the matching and of the parton-shower programs lead to differences which are considerably larger than those
  observed at fixed-order, even in the experimental fiducial region. 
  We conclude with some recommendations for experimental studies of vector-boson scattering processes.

>>>>

  Vector-boson scattering processes are of great importance for the
  current run-2 and future runs at design energy of the Large Hadron
  Collider. The presence of triple and quartic gauge couplings in the process
  gives access the gauge sector of the Standard Model and possible new-physics
  contributions there. 
  To test this hypothesis, a precise knowledge of the Standard
  Model contributions is necessary, 
  with a precision which at least matches the
  experimental uncertainties
  of existing and forthcoming measurements.
  In this article, we present a detailed study of the vector-boson
  scattering process with two positively-charged leptons and missing
  transverse momentum in the final state, mediated predominantly by
  same-sign production of two $\PW$ bosons with positive charge.
  In particular, we first carry out a systematic comparison of the various 
  approximations that are usually performed for this kind of process against the complete calculation, 
  at LO and NLO QCD accuracy. Such  a study is performed both 
  in the usual fiducial region used by experimental collaborations and in a more inclusive phase space,
  where the differences among the various approximations lead to more sizeable effects.
  Afterwards, we turn ourselves to predictions matched to parton showers, at LO and NLO: we show that on the
  one hand, the inclusion of NLO QCD corrections leads to more stable predictions, but on the other the details 
  of the matching and of the parton-shower programs lead to differences which are considerably larger than those
  observed at fixed-order, even in the experimental fiducial region. 
  We conclude with some recommendations for experimental studies of vector-boson scattering processes.

### 2018-02-15 Draft/definition.tex
further small changes

The scattering of two positively charged W bosons is proceeding at the LHC through the partonic process:
%
\begin{equation}
\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}.
\end{equation}

@@

The process can also be mediated via a gluon connecting the two quark lines while the W bosons are radiated off the quark lines.
These contributions are of order $\mathcal{O}{\left(\alphas^{2}\alpha^{4}\right)}$ and feature different kinematical behaviours than the EW contribution.
Nonetheless they share the same final state and therefore constitute an irreducible background.

@@

In experimental measurements, special VBS-cuts are designed to enhance the EW contribution over the QCD one.
These cuts are based on the fact that the two contributions have different kinematical behaviour.
The EW contribution is characterised by two jets with large rapidities as well as a large invariant mass.
The two W bosons are mostly produced centrally.
This is in contrast with the QCD contribution which favours jets in the central region.
Therefore, the event selection usually involves rapidity-difference and invariance-mass cuts for the jets.
This will be further discussed in Sec.~\ref{subsec:contributions}.

>>>>

The scattering of two positively charged $\PW$ bosons with their subsequent leptonic decay proceeds at the LHC through the partonic process:
%
\begin{equation}
\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}.
\end{equation}

@@

The process can also be mediated via a gluon connecting the two quark lines while the $\PW$ bosons are radiated off the quark lines.
These contributions are of order $\mathcal{O}{\left(\alphas^{2}\alpha^{4}\right)}$ and feature different kinematical behaviours than the EW contribution.
Nonetheless they share the same final state and therefore constitute an irreducible background.

@@

In experimental measurements, special VBS-cuts are designed to enhance the EW contribution over the QCD one.
These cuts are based on the fact that the two contributions have different kinematical behaviour.
The EW contribution is characterised by two jets with large rapidities as well as a large invariant mass.
The two $\PW$ bosons are mostly produced centrally.
This is in contrast with the QCD contribution which favours jets in the central region.
Therefore, the event selection usually involves rapidity-difference and invariant-mass cuts for the jets.
This will be further discussed in Sec.~\ref{subsec:contributions}.

### 2018-02-15 Draft/process.tex
some improvements in the draft
further small changes

As mentioned previously, the contribution of main interest in our process is
the scattering of two W gauge bosons, which includes the quartic gauge-boson vertex.
Therefore it is justified to approximate the full EW contributions simply by only these contributions which contain the $2\rightarrow 2$ scattering process as a subpart.
However, this set of contributions is not gauge invariant.
To make it gauge invariant, one should perform an on-shell projection of the incoming W bosons.
Unfortunately, these momenta are space-like and thus a simple on-shell projection is not possible.
Instead, one can keep the W boson legs connected to the external quark line off-shell while the ones connected to the final-state leptons, which are already time-like, are put on-shell.
Then the polarisation of the gauge boson is accommodated following for example the implementations of Refs.~\cite{Kuss:1995yv,Accomando:2006hq}.
Such an approximation is usually called effective vector-boson approximation (EVBA) \cite{Dawson:1984gx,Duncan:1985vj,Cahn:1983ip}.

A less crude approximation consists in considering all $t$- and $u$- diagrams and squaring them separately, neglecting interference contributions between the two.
These interferences are expected to be small in the VBS fiducial region, as they are both phase-space and colour suppressed.
The $s$-channel squared diagrams and any interferences with $s$-channels are left out as well.
This approximation is often called $t$-/$u$- approximation, VBF, or even VBS approximation.
We will adopt the latter denomination in the following of the article.
Such an approximation is implemented at LO in the computer codes {\sc Bonsay} and {\sc Powheg}.
This approximation is gauge-invariant, which can be seen by considering that the two protons and therefore the two incoming quarks belong to two different, but otherwise identical, copies of the $SU \left(3\right)$ gauge group.

The squared matrix element of the $s$-channel contributions can be added in addition, but all interferences between different kinematic channels are still neglected.
This is the level of simulation available in {\sc VBFNLO}.

All other codes ({\sc MG5\_aMC}, {\sc MoCaNLO+Recola}, {\sc PHANTOM}, and {\sc Whizard}) consider all contributions of order $\mathcal{O}{\left(\alpha^{6}\right)}$ as well as all possible interferences.
Note that the final W boson can always be considered either on-shell or off-shell without affecting the previous discussion.
All the codes mentioned here are described in details in the following sub-section.

@@

A further refinement is to consider the full real contributions as well as part of the virtual.
In particular one can consider only one-loop amplitudes where there is no gluon exchange between the quarks and assuming a cancellation of the infrared (IR) poles. \MP{True? I cannot remember exactly what is included}.
\MR{What about gauge invariance? A comment about that would be useful as well.}
Such predictions are provided by {\sc MG5\_aMC}.

>>>>

As mentioned previously, the contribution of main interest in our process is
the scattering of two $\PW$ gauge bosons, which includes the quartic gauge-boson vertex.
Therefore it is justified to approximate the full EW contributions simply by only these contributions which contain the $2\rightarrow 2$ scattering process as a subpart.
However, this set of contributions is not gauge invariant. In order to make it gauge invariant, one 
should perform an on-shell projection of the incoming and outgoing $\PW$ bosons {\bf MZ: it is incoming and outgoing, right? in case modify what follows}.
Unfortunately, these momenta are space-like and thus a simple on-shell projection is not possible.
Instead, one can keep the $\PW$ boson legs connected to the external quark line off-shell while the ones connected to the final-state leptons, which are already time-like, are put on-shell.
Then the polarisation of the gauge boson is accommodated following for example the implementations of Refs.~\cite{Kuss:1995yv,Accomando:2006hq}.
Such an approximation is usually called effective vector-boson approximation (EVBA) \cite{Dawson:1984gx,Duncan:1985vj,Cahn:1983ip}.

A less crude approximation consists in considering all $t$- and $u$- diagrams and squaring them separately, neglecting interference contributions between the two.
These interferences are expected to be small in the VBS fiducial region, as they are both phase-space and colour suppressed.
The $s$-channel squared diagrams and any interferences with $s$-channels are also discarded.
This approximation is often called $t$-/$u$- approximation, VBF, or even VBS approximation.
We will adopt the latter denomination in the following of the article.
Such an approximation is implemented at LO in the computer codes {\sc Bonsay} and {\sc Powheg}.
This approximation is gauge-invariant, which can be seen by considering that the two protons and therefore the two incoming quarks belong to two different, but otherwise identical, copies of the $SU \left(3\right)$ gauge group.

A further refinement is to add the squared matrix element of the $s$-channel contributions, still neglecting 
all interferences between different kinematic channels, as it is implemented in in {\sc VBFNLO}.

All other codes ({\sc MG5\_aMC}, {\sc MoCaNLO+Recola}, {\sc PHANTOM}, and {\sc Whizard}) consider all contributions of order $\mathcal{O}{\left(\alpha^{6}\right)}$ as well as all possible interferences.
Note that the final $\PW$ boson can always be considered either on-shell or off-shell without affecting the previous discussion.
All the codes mentioned here are described in details in the following sub-section.

@@

A further refinement is to consider the full real contributions as well as part of the virtual.
In particular one can consider only one-loop amplitudes where there is no gluon exchange between the quarks and 
assuming a cancellation of the infrared (IR) poles. Such an approximation is implemented by {\sc MG5\_aMC}.
